Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-07-06 paper_jasa.tex
tweaked abstract

\begin {abstract}  
Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis. Until recently there were no formal visual methods in place for determining statistical significance of findings. This changed, when \citet{buja:2009} conceptually introduced two protocols for formal tests of visual findings. In this paper we take this a step further by comparing the lineup protocol \citep{buja:2009} against conventional statistical testing  of the significance of regression model parameters. \blue{We extend and refine the terminology of visual inference to a more general framework to allow for comparisons with conventional tests.} A human subjects experiment is conducted using simulated data to provide controlled conditions. Results suggest that the lineup protocol provides results comparable to the conventional tests, out-performs them when data is contaminated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult trials.
\end {abstract}

>>>>

\begin {abstract}  
Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis. Until recently, there were no formal visual methods in place for determining statistical significance of findings. This changed when \citet{buja:2009} conceptually introduced the lineup protocol for formal tests of visual findings. In this paper this is taken a step further by refining the terminology of visual inference, framing the protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists, and it is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models. A human subjects experiment is conducted using simulated data to provide controlled conditions. Results suggest that the lineup protocol provides results comparable to the conventional tests, out-performs them when data is contaminated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult trials.
\end {abstract}
